MLH1 Results Outline

### 2020-02-14 doc/Results.rmd
trying to fix bug for kniting
knit successful, adding more results

load(file="~./MLH1repo/data/CleanBivData_12.2.20.RData")

@@

glmMLH1Var_M3 <- lm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)

MLH1cv_M3 <- lm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
#summary(MLH1cv_M4)
round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)

### Q12 var
MLH1var_M3_Q12 <- lm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )

@@

#Q12 cv
MLH1cv_M4_Q12 <- lm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)

#summary(MLH1cv_M4_Q12)
#anova(MLH1cv_M4_Q12)
#as.data.frame( round(summary(MLH1cv_M4_Q12)$coefficients, 6))

@@

```{r Q1_short_biv_lmer, echo=FALSE}

@@

```{r Q1.SC.length.show_long.short, echo=FALSE}

@@

Q1.SClength_Musc.BivData_plot

@@

```{r Q1show.F1pos_Full, echo=FALSE}

@@

```{r CO1_foci_pos_boxplot, echo=FALSE}

@@

### Siscoten

@@

### Heterochiasmy Prediction

> Is sex a significant effect for the 1CO normalized CO position? (as predicted)

> is random strtain effect significant?

The random strain effect seems very significant.

remember to use the mouse average table **mouse.avs_4MM**  (I don't think I need the MELT data frame)

```{r pos_t.test}
#make point plots / boxplots which show differences in mean positions

#scatter + boxplot for t.tests
```

Dr. Broman suggested that the Smirnov K /(curve comparison) wasn't the best test to differences in general CO position. He suggested doing simple t-test for the positions

```{r F1.by.F2}

```{r CO.pos.ANOVA, eval=FALSE}
#what is the pattern of variance
#run analyses for each subsp*sex
#use non-melt DF

#how is the variance partioned across
#cell, mouse, strain

female.Dom <- Curated_BivData[Curated_BivData$sex == "female",]
female.Dom <- female.Dom[female.Dom$subsp == "Dom",]

female.Dom$Foc1.PER <- female.Dom$Foci1 / female.Dom$chromosomeLength

#unorder strain and mouse

female.Dom$mouse <- as.factor(female.Dom$mouse)

female.Dom$strain <- unclass(female.Dom$strain)
female.Dom$strain <- as.factor(female.Dom$strain)

female.Dom_1CO <- female.Dom[female.Dom$hand.foci.count == 1,]
female.Dom_1CO <- female.Dom_1CO[(!is.na(female.Dom_1CO$hand.foci.count)),]

#1CO first
modo <- lm(Foc1.PER ~ fileName + mouse + strain, data=female.Dom_1CO)

#can't get mouse and strain to have sum of square
#residual size decreases with per.F1
#residuals much larger than fileName, mouse and strain no

#model <- lm(breaks ~ wool * tension, 
#            data = warpbreaks, 
#            contrasts = list(wool = "contr.sum", tension = "contr.poly"))

male.Dom <- Curated_BivData[Curated_BivData$sex == "male",]
male.Dom <- male.Dom[male.Dom$subsp == "Dom",]

male.Dom$mouse <- as.factor(male.Dom$mouse)

male.Dom$strain <- unclass(male.Dom$strain)
male.Dom$strain <- as.factor(male.Dom$strain)

male.Dom <- male.Dom[male.Dom$hand.foci.count == 1,]
male.Dom <- male.Dom[(!is.na(male.Dom$hand.foci.count)),]

male.Dom$Foc1.PER <- male.Dom$Foci1 / male.Dom$chromosomeLength

male.modo <- lm(Foc1.PER ~  fileName | mouse | strain, data=male.Dom)
summary(aov(male.modo))

#only file name is registering as effect
#Review ANOVA frameworks
#http://www.biostathandbook.com/nestedanova.html

@@

Brief lit review for IFD / interference expectations  
**Petkov et al 2007,  CO interference underlies sex differences in RR**

@@

Chrm1 genetic map,

(other human and mice refs

Tease n Hulten 2004: -- no difference between MLH1 foci in males and females

@@

@@

C.2) sis-co-ten metric ... (what about the clustering?)

C.3) telomere and centromere distances ...

M1 glm model for fixed strain effect across male averages.

@@

@@

```{r Q2.male.Nrm1CO_strain, echo=FALSE}
#isolate all musc
#Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]

@@

@@

```{r Q2.SC.long.biv_2, echo=FALSE}

#>>>>>>> 7ebba2e601079e1783e92c05ba74529ea765a5ed

shoo.long.biv <- ggplot(male.long.biv, aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")
shoo.long.biv

cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av. \nMale Long Biv SC Lengths")
cell.avlong.biv

mouse.shoo.long.biv <- ggplot(mouse.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Mouse av. \nMale Long Biv SC Lengths")
mouse.shoo.long.biv

cell.avlong.biv <- ggplot(cell.av.long.biv.table, aes(y=mean.sc, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Cell av. \nMale Long Biv SC Lengths")
cell.avlong.biv

Not the pattern I expected, there might be more noise than I thought in this dataset.

```{r Q2.SC.long.biv_3, echo=FALSE}

#mouse.av.long.biv.table <- make this is set up

glm1.SC.male_long.biv <- glm(mean_SC ~ subsp * strain, data= Long.biv_mouse_table)
#summary(glm1.SC.male_long.biv)

glm2.SC.male_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table)
#summary(glm2.SC.male_long.biv)

#Dom
glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
#summary(glm_SC.male_Dom_long.biv)  #ALL strain are sig
#both G and LEW have significant mean.SC, doesn't meet predictions

#Musc
glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
#summary(glm_SC.male_Musc_long.biv)

#Mol
glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
#summary(glm_SC.male_Mol_long.biv)

@@

@@

#### Q2 SC by class

@@

plot.blah.mouse.av <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
                          aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av. SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)
plot.blah.mouse.av

plot.logReg.mouse.av_1CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
                          aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av. SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
plot.logReg.mouse.av_1CO

plot.logReg.mouse.av_2CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
                          aes(x=mean.SC_2CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av. SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)
plot.logReg.mouse.av_2CO

@@

```{r Q2.SClength.t.test, echo=FALSE}
#MOUSE AVE, all
Rec.group.mouse.SC_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
                                 Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])

@@

### Q2 SC length Chrm class prediction

@@

>>>>

load(file="~./MLH1repo/data/CleanBivData_13.2.20.RData")

@@

glmMLH1Var_M3 <- glm( var ~ sex * strain,  data=DF.HetC.MixedModel.HQ )
glmMLH1Var_M3_sum <- summary(glmMLH1Var_M3)
round(glmMLH1Var_M3_sum$coefficients[,c(1,4)], 5)

MLH1cv_M3 <- glm(cV ~ sex * strain, data=DF.HetC.MixedModel.HQ)
#summary(MLH1cv_M4)
round(summary(MLH1cv_M3)$coefficients[,c(1,4)], 6)

### Q12 var
MLH1var_M3_Q12 <- glm( var ~ sex*strain, data=Q12_mouse_table_w.Ages )

@@

#Q12 cv
MLH1cv_M4_Q12 <- glm( cV ~ sex*strain, data=Q12_mouse_table_w.Ages )
round(summary(MLH1cv_M4_Q12)$coefficients[,c(1,4)], 6)

@@

```{r Q1_short_biv_lmer, echo=FALSE}
#Mouse.Table_BivData_4MM #this one is the HetC table (all the Curated BivData)
#Long.biv_mouse_table_HetC  (long biv data set for HetC models)

@@

```{r Q1.SC.length.show_long.short, echo=FALSE, eval=FALSE}

@@

@@

```{r Q1show.F1pos_Full, echo=FALSE, eval=FALSE}

@@

```{r CO1_foci_pos_boxplot, echo=FALSE, eval=FALSE}

@@

<!--
### Siscoten

@@

@@

<!-- 
Brief lit review for IFD / interference expectations  
**Petkov et al 2007,  CO interference underlies sex differences in RR**

@@

Chrm1 genetic map, 
(other human and mice refs 
Tease n Hulten 2004: -- no difference between MLH1 foci in males and females

@@

total.IFDs <- Curated_BivData %>% filter(hand.foci.count > 1)

@@

@@

In order to test if sc lengths (sc area) is a predictor of genome wide recombination rate variation, we compared /analyzed the average length of bivalents between the high and low recombining strains.

The average sc lengths increase with the number of MLH1 foci / chromosome classes for all strains, but grouping bivalent observations by chromosome class (foci number) is confonded by the different proportions across strains.

long.biv data annd short bivdata

@@

```{r Q2.male.Nrm1CO_strain, echo=FALSE, eval=FALSE}
#isolate all musc
#Curated_BivData_male <- Curated_BivData[Curated_BivData$sex == "male",]

@@

```{r wacky, echo=FALSE}

real.long.bivData$bin <- "long"
real.short.bivData$bin <- "short"

lng.shrt <- rbind(real.long.bivData[,c(1:2, 17,19)], real.short.bivData[,c(1:2, 17,19)])

uu <- ggplot(lng.shrt, aes(x=fileName, y=chromosomeLength) )  + geom_point()+facet_wrap(~strain, scales="free")

```{r Q2.SC.short.long.setup, echo=FALSE}
table(male.long.biv$strain)
#the Long biv dataset is quite unbalance (esp MSM)
male.long.biv <- real.long.bivData %>% filter(sex== "male")

                        n_obs = length(Obj.ID),
                         mean.SC = mean(chromosomeLength)
                         )
cell.av.long.biv.table <- add_mouse(cell.av.long.biv.table)
cell.av.long.biv.table <- add_strain(cell.av.long.biv.table)
cell.av.long.biv.table <- add_subsp(cell.av.long.biv.table)

cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)
cell.av.long.biv.table$strain <- factor(cell.av.long.biv.table$strain, ordered = FALSE)

mouse.av.long.biv.table <- ddply(male.long.biv, c("mouse"), summarise,
                        n_obs = length(Obj.ID),
                         mean.SC = mean(chromosomeLength)
                         )
mouse.av.long.biv.table <- add_mouse(mouse.av.long.biv.table)
mouse.av.long.biv.table <- add_strain(mouse.av.long.biv.table)
mouse.av.long.biv.table <- add_subsp(mouse.av.long.biv.table)

mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)
mouse.av.long.biv.table$strain <- factor(mouse.av.long.biv.table$strain, ordered = FALSE)

male.short.biv <- real.short.bivData %>% filter(sex== "male")

mouse.av.short.biv.table <- ddply(male.short.biv, c("fileName"), summarise,

                        n_obs = length(Obj.ID),
                         mean.SC = mean(chromosomeLength)
                         )

mouse.av.short.biv.table <- add_mouse(mouse.av.short.biv.table)
mouse.av.short.biv.table <- add_strain(mouse.av.short.biv.table)
mouse.av.short.biv.table <- add_subsp(mouse.av.short.biv.table)
mouse.av.short.biv.table$strain <- factor(mouse.av.short.biv.table$strain, ordered = FALSE)
mouse.av.short.biv.table$strain <- factor(mouse.av.short.biv.table$strain, ordered = FALSE)
mouse.av.short.biv.table$strain <- factor(mouse.av.short.biv.table$strain, ordered = FALSE)
```

```{r Q2.SC.long.biv_2, echo=FALSE}
shoo.long.biv <- ggplot(male.long.biv, aes(y=chromosomeLength, x=strain, color=strain))+geom_jitter()+facet_wrap(~subsp, scales = "free")+scale_color_manual(values=colors_of_strains)+ylim(c(75,140))+geom_boxplot(alpha=.2) + ggtitle("Single Biv \nMale Long Biv SC Lengths")

@@

```{r Q2.glm.SC.long, echo=FALSE}
glm.Q2.M2.SC.cell.av_long.biv <- glm(mean.SC ~ strain, data= cell.av.long.biv.table)
round(summary(glm.Q2.M2.SC.cell.av_long.biv)$coefficients[,c(1,4)], 6)

table(mouse.av.long.biv.table$mouse)

glmQ2.M2.SC.male_long.biv <- glm(mean.SC ~ strain, data= mouse.av.long.biv.table)
round(summary(glmQ2.M2.SC.male_long.biv)$coefficients[,c(1,4)], 6)
```
In the cell average Long biv data set, MSM G and CZECH strains have sig strain effects

very few strains have significant values for the mouse

```{r Q2.glm.SC.short, echo=FALSE}
glmQ2_SC_mouse.av.short_biv <- glm(mean.SC ~ strain, data= mouse.av.short.biv.table)

round(summary(glmQ2_SC_mouse.av.short_biv)$coefficients[,c(1,4)], 6)
```
Basically all strains have a significant effect for the mouse.av of long biv

```{r Q2.glm.SC.all.strain, echo=FALSE}

#Dom
glm_SC.male_Dom_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Dom",])
round(summary(glm_SC.male_Dom_long.biv)$coefficients[,c(1,4)], 6)

#Musc
glm_SC.male_Musc_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Musc",])
round(summary(glm_SC.male_Musc_long.biv)$coefficients[,c(1,4)], 6)

#summary(glm_SC.male_Musc_long.biv)

#Mol
glm_SC.male_Mol_long.biv <- glm(mean_SC ~ strain, data= Long.biv_mouse_table[Long.biv_mouse_table$subsp == "Mol",])
round(summary(glm_SC.male_Mol_long.biv)$coefficients[,c(1,4)], 6)

@@

abova are glm tables for mouse.av SC lengths for each strain.

@@

@@

plot.blah.mouse.av <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
                          aes(x=mean_SC, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av. SC (pooled) \nHigh vs low gwRR strains")+facet_wrap(~subsp)

plot.logReg.mouse.av_1CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
                          aes(x=mean.SC_1CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av. SC 1CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)

plot.logReg.mouse.av_2CO <- ggplot(Male.poly.Mouse.Table_BivData_4MM,
                          aes(x=mean.SC_2CO, y=Rec.group)) + geom_point(aes(color=strain)) +  scale_color_manual(values=colors_of_strains)+ stat_smooth(method="glm", method.args=list(family="binomial"), se=FALSE) + ggtitle("Mouse av. SC 2CO \nHigh vs low gwRR strains")+facet_wrap(~subsp)

@@

```{r Q2.SClength.ttest, echo=FALSE}
#MOUSE AVE, all
Rec.group.mouse.SC_Ttest <- t.test(Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 1], 
                                 Male.poly.Mouse.Table_BivData_4MM$mean_SC[Male.poly.Mouse.Table_BivData_4MM$Rec.group == 0])

@@

@@

```{r F1.by.F2, eval=FALSE}
